Dismal Code: Studying the Evolution of Security Bugs

### 2013-02-12 docs/msr.tex
table captions should be on top
added graph results description

\begin{itemize}
	\item The analysis of how security bugs evolve over time. To achieve
this, we inspect every project per version. Our hypothesis is that security
bugs must decrease as a project evolves since they are critical and developers
should eliminate them.
	\item We show evidence of the relation between security bugs and the artifact's size.  Here, size is
defined in terms of bytecode.
%	\item Security bug persistense between versions. We show evidence that specific bugs are
%not eliminated between versions.
	\item The correlation of security bugs with other bug categories. We
show evidence that such bugs are correlated with bugs that have to do with
performance, coding practices and others.
	\item The relation of software dependencies and security bugs. Our
hypothesis is that an artifact that is a dependency of many other artifacts
should have a low rate of security bugs. 
\end{itemize}

@@

Our experiment goal was to collect the metric results of the FindBugs tool.
The process involved four entities: a number of {\it workers}, a
{\it task queue} mechanism,\footnote{\url{http://www.rabbitmq.com/}}
a {\it data repository},\footnote{\url{http://www.mongodb.org/}}
and the {\it code repository}, which in our case it was the public Maven 
repository\footnote{\url{http://mvnrepository.com}}.

\begin{table}
\centering
\caption{Experiment software and versions}
\label{tbl:soft-version}
\begin{tabular}{r r r}
\hline
 & Software Package & version\\
 \hline
worker & Custom Python Script & 2.7\\
task queue & RabbitMQ\footnote{\url{http://www.rabbitmq.com/}} & \\
data repository & MongoDB\footnote{\url{http://www.mongodb.org/}} & \\
code repository & Maven Repository & \\
\hline
\end{tabular}
\end{table}

@@

Another observation involves bugs that we could call {\it
severe}. Such bugs are related to vulnerabilities that appear due to the lack of user-input
validation and can lead to damaging attacks like {\sc sql} injection and
Cross-Site Scripting. Note that to exploit such vulnerabilities, a malicious user does
not have to know anything about the applications internals. For all the other
bugs, another program should be written either to incorporate references to
mutable objects, access non-final fields and others. In addition, as
bug descriptions indicate,\footnote{\url{http://findbugs.sourceforge.net/bugDescriptions.html}}
if an application has bugs like these, it might have more vulnerabilities that FindBugs doesn't
report. Table~\ref{tbl:sev} presents the number of artifacts where at least one of these
bugs exists. In the following section we provide results separately for this
subcategory.

\begin{table*}
\centering
\begin{center}
\leavevmode
	\begin{tabular}{l r}
	\hline
	Bug Description & Number of Artifacts\\
 	\hline
	{\sc hrs}: {\sc http} cookie formed from untrusted input & 149\\
	{\sc hrs}: {\sc http} response splitting vulnerability & 1,564\\
	{\sc pt}: absolute path traversal in servlet  & 92\\
	{\sc pt}: relative path traversal in servlet & 50\\
	{\sc sql}: nonconstant string passed to execute method on an {\sc sql} & 1,817\\
	{\sc sql}: a prepared statement is generated from a nonconstant String & 1,438\\
	{\sc xss}: {\sc jsp} reflected cross site scripting vulnerability & 17\\
	{\sc xss}: Servlet reflected cross site scripting vulnerability in error page & 88\\
	{\sc xss}: Servlet reflected cross site scripting vulnerability & 140\\
	\hline
	\end{tabular}
\caption{Number of Artifacts that Contain at Least One Severe Security Bug}
\label{tbl:sev}
\end{center}
\end{table*}

@@

@@

>>>>

\begin{itemize}
	\item The analysis of how security bugs evolve over time. To achieve
this, we inspect every project per version. Our hypothesis is that security
bugs must decrease as a project evolves since they are critical and developers
should eliminate them.
	\item We show evidence of the relation between security bugs and the artifact's size.  Here, size is
defined in terms of bytecode.
	\item Security bug persistense between versions. We show evidence that specific bugs are
not eliminated between versions.
	\item The correlation of security bugs with other bug categories. We
show evidence that such bugs are correlated with bugs that have to do with
performance, coding practices and others.
	\item The relation of software dependencies and security bugs. Our
hypothesis is that an artifact that is a dependency of many other artifacts
should have a low rate of security bugs. 
\end{itemize}

@@

Our experiment goal was to collect the metric results of the FindBugs tool.
The process involved four entities (see Table~\ref{tbl:exp}):
a number of {\it workers}, a {\it task queue}
mechanism,\footnote{\url{http://www.rabbitmq.com/}}
a {\it data repository},\footnote{\url{http://www.mongodb.org/}}
and the {\it code repository}, which in our case it was the public Maven 
repository\footnote{\url{http://mvnrepository.com}}.

\begin{table}
\centering
\caption{Experiment Entities}
\label{tbl:exp}
\begin{tabular}{r r r}
\hline
 & Software Package & version\\
 \hline
Worker & Custom Python Script & 2.7\\
Task Queue & RabbitMQ & 3.0.1 \\
Data Repository & MongoDB & 2.2 \\
Code Repository & Maven Repository & January 2012 \\
\hline
\end{tabular}
\end{table}

@@

Another observation involves bugs that we could call {\it
severe}. Such bugs are related to vulnerabilities that appear due to the lack of user-input
validation and can lead to damaging attacks like {\sc sql} injection and
Cross-Site Scripting. Note that to exploit such vulnerabilities, a malicious user does
not have to know anything about the applications internals. For all the other
bugs, another program should be written either to incorporate references to
mutable objects, access non-final fields and others. In addition, as
bug descriptions indicate,\footnote{\url{http://findbugs.sourceforge.net/bugDescriptions.html}}
if an application has bugs like these, it might have more vulnerabilities that FindBugs doesn't
report. Table~\ref{tbl:sev} presents the number of {\sc jar}s where at least one of these
bugs exists. In the following section we provide results separately for this
subcategory.

\begin{table*}
\centering
\caption{Number of Artifacts that Contain at Least One Severe Security Bug}
\label{tbl:sev}
\leavevmode
	\begin{tabular}{l r}
	\hline
	Bug Description & Number of Artifacts\\
 	\hline
	{\sc hrs}: {\sc http} cookie formed from untrusted input & 149\\
	{\sc hrs}: {\sc http} response splitting vulnerability & 1,564\\
	{\sc pt}: absolute path traversal in servlet  & 92\\
	{\sc pt}: relative path traversal in servlet & 50\\
	{\sc sql}: nonconstant string passed to execute method on an {\sc sql} & 1,817\\
	{\sc sql}: a prepared statement is generated from a nonconstant String & 1,438\\
	{\sc xss}: {\sc jsp} reflected cross site scripting vulnerability & 17\\
	{\sc xss}: Servlet reflected cross site scripting vulnerability in error page & 88\\
	{\sc xss}: Servlet reflected cross site scripting vulnerability & 140\\
	\hline
	\end{tabular}
\end{table*}

@@

As we mentioned earlier, during the experiment we managed to retrieve the
dependencies of every version. Based on this information we created a graph
that represented the snapshot of the repository that we were analyzing. The
nodes of the graph represented the versions and the vertices their dependencies.
We have to mention here that the graph is not representative. For instance, if
a dependency was pointing only to a project (and not to a specific version), we chose to
select the latest version found on the repository. Also, this graph is not
complete. This is because threre were missing versions.
Practically, the corresponding {\it pom}, or {\it war} files existed on the
repository, but the {\sc jar} file did not. From the 56,5680 vertices, 191,433
belonged to the first category while 16,4234 belonged to the second. Finally,
the graph contained 80,354 nodes. Obviously the number does not correspond to
the number of the total versions (see Section~\ref{sec:data}). This is because
some versions did not contain any information about their dependencies so they
are not represented in the graph. After creating the graph, we ran the PageRank
algorithm~\cite{BP98} on it and retrieved all PageRanks for every node. Then we
examined the security bugs of the 50 most popular nodes based on their PageRank.
33 of them contained security bugs while two of them contained severe bugs.
Finally, 25 of them were latest versions at the time.

@@

\section{Discussion}
\label{sec:dis}
